SDC: Social Dispersed Computing

### 2017-06-22 papers/IARPA-RFI-17-03/rfi.tex
fixes

The classic Internet architecture assumes computing hosts whose task is to perform computations
and a network of switches and routers whose sole task is to route network packets. This architecture delegates all computing to end-nodes and all communications to the network. However every resource (node or edge) in this architecture has a finite capacity potentially leading to inefficient allocation of resources to communications and computations.  This is the networking model on which our current state of the art cloud computing architectures are built. While this is is sufficient for most tasks it does not scale up when we try to use this architecture in the new emerging pardigm of Digitalization where the boundaries between the network node, the sensor, the actuator are blurring, driven primarily by the internet of things and the smart phone revolution. The data generated via this network is extremely large and needs clever, scalable and possibly decentralized computing solutions that can scale indepently as required. In this network any node in the graph can serve as a computing or network
router node (or both), and  the complex application can possibly be distributed over the network
of these nodes, such that the overall performance (e.g., amount of data processed over
time) is significantly improved.  This is the concept behin the `Social Dispersed Computing', which our groupt has been working on. In this white paper, we describe two usecases for this new computing paradigm and describe the challenges which must be solved.

@@

Fault tolerance and adaptivity to changing resources are important requirements for many systems. Our approach, through the underlying framework provides a solution so that complex applications can operate autonomously and efficiently, even under adversary conditions. Both fault tolerance and adaptivity require cooperation from the developers - i.e. the application framework cannot solve the problem, but the framework will provide services to help achieve the required system level fault-tolerance and adaptivity.  However, there are additional challenges, which we will describe using two case studies. The first one looks at the transactive energy pardigm, the other looks at multi-modal transporation analytics and optimization, a critical concept that is required to solve the large transportation challenges that we are facing.

@@

Urban spaces are expanding at impressive rates and an efficient transportation system is a key component to any well-functioning city. Yet, as cities expand, existing infrastructures are being stressed to their limits. To make the problem worse,  building new infrastructures is costly, time consuming, and disruptive. Consequently, a number of transit agencies and cities are turning towards multimodal  solutions combining transporation operated by both cities and private users. However, given the large number of choices and available options this often leads to a dynamic programming problem that must be solved online considering serveral state variables, such as traffic congenstion, mode capacity and the personal preferences.

@@

However, this precisely leads to the problem of secure and trustworthy computing.  Privacy of individuals is an important aspect of this solution; use of smart devices of individuals as both a data source and a computational resource  could have the effect of exposing the end-user to risk of a privacy breach. Seemingy innocuous data such as transit mode or route choice can lead to inferences of
private information such as real-time tracking of an individual's position~\cite{koufogiannis:2015aa}, likelihood of affairs~\cite{mueffelmann:2015aa}, forecasting trip destinations~\cite{dewri:2013aa}, etc.

>>>>

The classic Internet architecture assumes computing hosts whose task is to perform computations
and a network of switches and routers whose sole task is to route network packets. This architecture delegates all computing to end-nodes and all communications to the network. However, every resource (node or edge) in this architecture has a finite capacity potentially leading to inefficient allocation of resources to communications and computations.  This is the networking model on which our current state-of-the-art cloud computing architectures are built. While this is sufficient for most tasks, it does not scale up when we try to use this architecture in the new emerging paradigm of Digitalization. In this new paradigm, the boundaries between the network node, the sensor, and the actuator are blurring, driven primarily by the Internet of Things and the smartphone revolution. The data generated via this network is extremely large and needs clever, scalable, and possibly decentralized computing solutions that can scale independently as required. In this network, any node in the graph can serve as a computing or network
router node (or both), and  the complex application can possibly be distributed over the network
of these nodes, so that the overall performance (e.g., amount of data processed over
time) is significantly improved.  This is the concept behind `Social Dispersed Computing,' on which our group has been working. In this whitepaper, we describe two use-cases for this new computing paradigm and describe the challenges which must be solved.

@@

Fault tolerance and adaptivity to changing resources are important requirements for many systems. Our approach, through the underlying framework provides a solution so that complex applications can operate autonomously and efficiently, even under adversary conditions. Both fault tolerance and adaptivity require cooperation from the developers - i.e. the application framework cannot solve the problem, but the framework will provide services to help achieve the required system level fault-tolerance and adaptivity.  However, there are additional challenges, which we will describe using two case studies. The first one looks at the transactive energy paradigm, the other looks at multi-modal transportation analytics and optimization, a critical concept that is required to solve the large transportation challenges that we are facing.

@@

Urban spaces are expanding at impressive rates and an efficient transportation system is a key component to any well-functioning city. Yet, as cities expand, existing infrastructures are being stressed to their limits. To make the problem worse,  building new infrastructures is costly, time consuming, and disruptive. Consequently, a number of transit agencies and cities are turning towards multimodal  solutions combining transportation operated by both cities and private users. However, given the large number of choices and available options this often leads to a dynamic programming problem that must be solved online considering several state variables, such as traffic congestion, mode capacity and the personal preferences.

@@

However, this precisely leads to the problem of secure and trustworthy computing.  Privacy of individuals is an important aspect of this solution; use of smart devices of individuals as both a data source and a computational resource  could have the effect of exposing the end-user to risk of a privacy breach. Seemingly innocuous data such as transit mode or route choice can lead to inferences of
private information such as real-time tracking of an individual's position~\cite{koufogiannis:2015aa}, likelihood of affairs~\cite{mueffelmann:2015aa}, forecasting trip destinations~\cite{dewri:2013aa}, etc.
